### 2017-01-20 Risk analysis/Text/Layout/Layout/MSc_layout.tex

%% THIS IS THE MAIN FILE %%
% Here the main arrangement of your thesis is determined. 
% In this file you read-in all chapters (as separate .tex files)
% This is also the file you 'Build' with a LaTeX editor, like TeXmaker.
% The .cls file (MScThesis.cls) contains all details regarding copyrights and so on. Take a look and adjust where applicable. But be careful changing this file: it determines the whole lay-out of your thesis!
% This MSc thesis standard layout is optimized for double sided printing on A4 format paper.
% Good luck and have fun with your Master Research in Applied Geophysics! Kind regards, Niels Grobbe

\documentclass[a4paper,11pt]{MScThesis}
%
%\usepackage{pifont}
\mscName{Fabian Antonio Stamm}
\mscDate{\today}
\mscTitle{Test}
\mscSubTitle{Subtitle}
\mscKeyWords{thesis, msc, subject}
%
%\mscBackPicture{2660PF3}    % eps of 21 * 29.7 cm
\mscReaderOne{Prof. Florian Wellmann, Ph.D.}
\mscReaderTwo{Prof. Dr. Janos Urai}
\mscReaderThree{Miguel de la Varga, M.Sc.}
%\mscReaderFour{}
%
\setThesisInfo
%
\begin{document}
%
%============================= Front matter ========================================
\frontmatter %
%
% Make a hell of a lot of title pages
    \maketitle
%
% Abstract
    \nonumchap{Abstract}

Please pay particular attention to the preparation of your abstract; use this text as a guide. Every master thesis report must be accompanied by an informative abstract of no more than one paragraph (max 300 words). The abstract should be self-contained. No references, figures, tables, or equations are allowed in an abstract. Do not use new terminology in an abstract unless it is defined or is well-known from the literature. The abstract must not simply list the topics covered in the paper but should (1) state the scope and principal objectives of the research, (2) describe the methods used, (3) summarize the results, and (4) state the principal conclusions. Do not refer to the master thesis report itself in the abstract. For example, do not say, "In this thesis we will discuss". Furthermore the abstract must stand alone as a very short version of the master thesis report rather than as a description of the contents. Remember that the abstract will be the first and most widely read portion of the master thesis report. Readers will be influenced by the abstract to the point that they decide to read the master thesis report or not.

    \cleardoublepage
%
% Acknowledgements
    \nonumchap{Acknowledgements}%
    First of all I want to thank all the people who have participated in this project ..
    Remember, often more people have (in some way) contributed to your final thesis than you would initially think of....
    \vspace*{15mm}

    \noindent
    Delft University of Technology \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
    Swiss Federal Institute of Technology \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
    RWTH Aachen University \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
    \mscdate

%
% table of contents, (\toc or \toclof or \tocloflot )
    \tocloflot
%
% Nomenclature
    \printnomencl %
%
% Acronyms
    \nonumchap{Acronyms} %
    %
    \begin{acronym}%
        \acro{DUT}{Delft University of Technology}%
        \acro{ETH}{Swiss Federal Institute of Technology}%
        \acro{RWTH}{Aachen University}%
    \end{acronym}%
    %
    \cleardoublepage%
%
%
%============================= Main matter =========================================
%
\mainmatter
%
% Introduction
\include{chapter1} % notice how the reference to the introduction takes place.. It refers to the name.tex

\cleardoublepage

%
% First Part
    \part{First Part} % you can divide your thesis into different parts, for example Theory&Modeling in part 1 and real data examples in part 2.
		\include{chapter2}

%
% Bibliography
    \bibliographystyle{apalike}
    \printbib{MyBib} % in MyBib.bib you add all your reference information, following the correct format.
    % The *.bin file needs to be compiled using the Bib-compile option of your LaTeX text editor. Sometimes, the bib file needs to be built several times, as well as the main file, before all references occur correctly in your PDF.

%============================= Back matter =========================================
\appendix

    \chapter{The back of the thesis}

    \section{An appendix section}

    \subsection{An appendix subsection with C++ Lisitng}

    \lstset{language=C++}
    \lstinputlisting{test.c}

    \subsection{A \matlab $ $ Listing}

    \lstset{language=matlab}
    \lstinputlisting{test.m}

    \chapter{Yet another appendix}

    \section{Another test section}

    Ok, all is well.

% Index
    \printindex
    \cleardoublepage

\end{document}