### 2017-08-09 Risk analysis/Text/Layout/Layout/chapter4_3Dcase.tex
Chaos - nothing is working

\chapter{Three-dimensional reservoir case}

	\section{Constructing the three-dimensional structural geological model}
	- place layer positional points in three lines, resembling three hypothetical seismic lines as base information
	-

	\section{Identifying trap structures in the geological model}