### 2017-09-04 Risk analysis/Text/Layout/Layout/chapter1.tex
Text Methods Progress

\chapter{Introduction} \label{chap::intro}

    - mention developments in geological modeling

    - economic significance of geological modeling: examples of application in petroleum, mining, aquifers, CCS, etc.

    - Bayesian analysis developments

    - decision-making in for such economic scenarios and projects which involve geosciences

    - classical approach (decision trees) vs. Bayesian approach, expression of decision-maker's interests in continuous loss functions

    - influence of uncertainty and work done by Wellmann & DeLaVarga