### 2017-11-24 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Abstract (needs revision)

%% THIS IS THE MAIN FILE %%
% Here the main arrangement of your thesis is determined. 
% In this file you read-in all chapters (as separate .tex files)
% This is also the file you 'Build' with a LaTeX editor, like TeXmaker.
% The .cls file (MScThesis.cls) contains all details regarding copyrights and so on. Take a look and adjust where applicable. But be careful changing this file: it determines the whole lay-out of your thesis!
% This MSc thesis standard layout is optimized for double sided printing on A4 format paper.
% Good luck and have fun with your Master Research in Applied Geophysics! Kind regards, Niels Grobbe

\documentclass[a4paper,11pt]{MScThesis}
%
%\usepackage{pifont}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{booktabs}
\graphicspath{Figures/}
%\usepackage{mathpazo}
\mscName{Fabian Antonio Stamm}
\mscDate{\today}
\mscTitle{Bayesian Decision Theory in Structural Geological Modeling}
\mscSubTitle{How Reducing Uncertainties Affects Reservoir Value Estimations}
\mscKeyWords{thesis, msc, subject}
%
%\mscBackPicture{2660PF3}    % eps of 21 * 29.7 cm
\mscReaderOne{Prof. Florian Wellmann, Ph.D.}
\mscReaderTwo{Prof. Dr. Janos Urai}
\mscReaderThree{Miguel de la Varga, M.Sc.}
%\mscReaderFour{}
%
\setThesisInfo
%
\begin{document}
%
%============================= Front matter ========================================
\frontmatter %
%
% Make a hell of a lot of title pages
    \maketitle
%
% Abstract
	\nonumchap{Abstract}
Bayesian decision theory is applicable to a wide spectrum of problems. Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision making is indispensable. Actors in this field are faced by numerous uncertainties that have to be considered. Structural geological modeling is of central importance for the assessment of uncertain hydrocarbon accumulations in potential reservoirs. By regarding such modeling as a Bayesian inference problem, additional geological information can be incorporated as likelihood functions linked to prior parameters in a probabilistic framework. Markov chain Monte Carlo sampling is used to approximate posterior models of reduced uncertainty. Here, synthetic geological models are constructed to represent potential hydrocarbon system. Algorithms for automatic trap recognition are developed to enable the reservoir valuation of prior and posterior models. These serve as a base for true value estimation by respective optimization of a case-specific loss function. This function is customized to reflect the decision making of differently risk-affine actors in this context. Results from conducting this for several inference cases show that the various Bayes estimators shift according to the characteristics of the underlying value distribution. While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss. These degree of these processes is considered a measure for the state of knowledge and decision uncertainty at the decision-making moment. This decision uncertainty does not change aligned with model uncertainty, but depends on alterations of critical parameter and interdependencies, relating to seal reliability in particular. Additionally, actors affected differently by one set of information, as defined by their risk-opportunity weighting. For good decision making, it appears to be most important to recognize in for which parameters reduction of uncertainty leads to changes in decisions.

%Please pay particular attention to the preparation of your abstract; use this text as a guide. Every master thesis report must be accompanied by an informative abstract of no more than one paragraph (max 300 words). The abstract should be self-contained. No references, figures, tables, or equations are allowed in an abstract. Do not use new terminology in an abstract unless it is defined or is well-known from the literature. The abstract must not simply list the topics covered in the paper but should (1) state the scope and principal objectives of the research, (2) describe the methods used, (3) summarize the results, and (4) state the principal conclusions. Do not refer to the master thesis report itself in the abstract. For example, do not say, "In this thesis we will discuss". Furthermore the abstract must stand alone as a very short version of the master thesis report rather than as a description of the contents. Remember that the abstract will be the first and most widely read portion of the master thesis report. Readers will be influenced by the abstract to the point that they decide to read the master thesis report or not.

    \cleardoublepage
%
% Acknowledgements
    \nonumchap{Acknowledgements}%
    First of all I want to thank all the people who have participated in this project ..
    Remember, often more people have (in some way) contributed to your final thesis than you would initially think of....
    \vspace*{15mm}

    \noindent 
    RWTH Aachen University \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
    \mscdate

%
% table of contents, (\toc or \toclof or \tocloflot )
    \tocloflot
%
% Nomenclature
    \printnomencl %
%
% Acronyms
    \nonumchap{Acronyms} %
    %
    \begin{acronym}%
        \acro{RWTH}{Aachen University}%
    \end{acronym}%
    %
    \cleardoublepage%
%
%
%============================= Main matter =========================================
%
\mainmatter
%
% Introduction
%\include{chapter1} % notice how the reference to the introduction takes place.. It refers to the name.tex

\cleardoublepage
		\include{chapter1}
		\include{chapter2_methods}
		\include{chapter3_results}
		\include{chapter4_discussion}
		%\include{chapter5_conclusion}
		%\include{chapter3_1Dcase}
		%\include{chapter4_3Dcase}

%
% Bibliography
    \bibliographystyle{apalike}
    \printbib{complete_biblio} % in MyBib.bib you add all your reference information, following the correct format.
    % The *.bin file needs to be compiled using the Bib-compile option of your LaTeX text editor. Sometimes, the bib file needs to be built several times, as well as the main file, before all references occur correctly in your PDF.

%============================= Back matter =========================================
\appendix

    \chapter{The back of the thesis}

    \section{An appendix section}

    \begin{figure}[h]
    	\centering
    	\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input1}
    	\caption{....}\label{fig:CSV_input1}
    \end{figure}

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input2}
		\caption{....}\label{fig:CSV_input2}
	\end{figure}

	\begin{figure}[h]
		\centering
		\includegraphics[width=1\textwidth]{Figures/Appendix/CSV_input3}
		\caption{....}\label{fig:CSV_input3}
	\end{figure}

    \begin{figure}[p!]
    	\centering
    	\includegraphics[width=1\textwidth]{Figures/Appendix/ML3}
    	\caption{....}\label{fig:ML3}
    \end{figure}

    \subsection{An appendix subsection with C++ Lisitng}

    \lstset{language=C++}
    \lstinputlisting{test.c}

    \subsection{A \matlab $ $ Listing}

    \lstset{language=matlab}
    \lstinputlisting{test.m}

    \chapter{Yet another appendix}

    \section{Another test section}

    Ok, all is well.

% Index
    \printindex
    \cleardoublepage

\end{document}